Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-27 ms.tex
cutting fat - and lots of it4

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  Thin solid and dot-dashed curves
    correspond to our analytic estimates of $\dot{M}/\dot{M}_{\rm
      edd}$ (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively. Equation ~[\ref{eq:eddr_analytic}] is derived from
    the approximate analytic result for the stagnation radius given by
    equation~\eqref{eq:stag_simple}. If we instead use the more
    accurate result from equation~\eqref{eq:rs2main} we obtain the
    thick curves.}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{cooling.eps}
  \caption{\label{fig:cooling} Ratio of the rate of heating to
radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$, as a
function of radius (solid lines) for each of our solutions shown in
Figure \ref{fig:profiles}.  Dashed lines show for comparison the ratio
of the cooling to the free-fall timescale $t_{\rm cool}/t_{\rm ff}$.
For high heating rates both ratios are approximately equal at the
stagnation radius (squares).  When $\dot{q}_{\rm
heat}/\dot{q}_{\rm rad} \lesssim 10$ (or, equivalently, $t_{\rm cool}/t_{\rm ff}
\lesssim 10$ near the stagnation radius), then the flow is susceptible to thermal instabilities.}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{vwSourcesImp.pdf}
\caption{\label{fig:vwSourcesImp} Different sources contributing to
  the gas heating rate in galactic nuclei at time $\tau_{\star}$ after
  a single burst of star formation.  Top and bottom panels show cases
  for black hole masses of $\Mbh=10^6 \Msun$ and $\Mbh= 10^8 \Msun$
  (both cusps with $\Gamma=0.8$).  Solid and dashed lines show the ranges of
  $\tau_{\star}$ for which the accretion flow is thermally stable and
  unstable, respectively, according to the ratio of $\dot{q}_{\rm
    heat}/|\dot{q}_{\rm rad}|$ near the stagnation radius
  (eq.~[\ref{eq:cooling3}]).  Also shown with horizontal gray lines
  are the values of the stellar velocity dispersion, $\sigma_0$,
  approximated as $\sqrt{3} \sigma_{\bullet}$, where
  $\sigma_{\bullet}$ is the velocity dispersion from the $\Mbh-\sigma$
  relation \citep{McConnellMa+:2011a}.}
\end{figure}

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of various
heating sources as a function of time $\tau_{\star}$ after a burst of
star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top
panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass
input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$
and $\eta(\tau_{\star})$, are calculated as described in
Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from
SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\rm C}$) are
calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC}
respectively, also making use of $\eta(\tau_{\star})$ from stellar
winds.  SN Ia heating depends also on a convolution of the DTD
(eq. [\ref{eq:DTD}]) and the SFH, which reduces to
the DTD itself for a single star burst.

We crudely account for the expected suppression of SN Ia heating
resulting from its non-steady nature by setting $v_{w}^{\rm Ia} = 0$
when the stagnation radius (calculated excluding Ia heating) exceeds
the Ia radius $r_{\rm Ia}$ exterior to which the interval between
subsequent supernovae exceeds the local dynamical time
(eq.~[\ref{eq:rIa}]).  Finally, note that $v_{\rm w}^{\rm C}$ and
$v_{\rm w}^{\rm Ia}$ depends on the total heating rate
(eq.~[\ref{eq:vtot}]), requiring us to simultaneously solve a series
of implicit equations to determine each heating rate.  Heating from
MSPs is found to be negligibly small for fiducial parameters and hence
is not shown, while kinetic feedback is neglected given its uncertain
efficiency ($\S\ref{sec:kinetic}$).

For low-mass black holes (Fig.~\ref{fig:vwSourcesImp}, top panel),
stellar winds are the most important source of heating on timescales
of $\tau_{\star} \lesssim 10^{7}$ years.  Compton heating comes to
dominate at later times due to the higher accretion rates that
accompany the overall decrease in the heating rate, coupled with the
persistently high mass loss rates and gas densities associated with a
relatively young stellar population.  Interestingly, Ia SN dominate
the heating briefly around 10$^{9}$ yr due to a window of time in
which $r_{ \rm Ia} < \rs$.  For high-mass black holes
(Fig.~\ref{fig:vwSourcesImp}, bottom panel), stellar wind heating
dominates early epochs, while SN Ia dominate after $10^{7}$ yr.

As the total heating rate declines with time, after some point the
flow becomes thermally unstable according to the criterion
$(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim 10$
(eq.~[\ref{eq:cooling3}]), as shown by dashed lines in
Fig.~\ref{fig:vwSourcesImp}.  Compton heating is neglected in
calculating thermal stability because, unlike local stellar feedback
mechanisms, SMBH feedback may not be capable of truly stabilizing the
flow, given its inability to respond instantaneously to local changes
in gas properties.  Thermally unstable flow is present
throughout the single starburst case except at very early times,
$\tau_{\star} \lesssim 10^7$ yr, and at late times $\tau_{\star}
\gtrsim 10^{9}$ yr in the high $M_{\bullet}$ case.

>>>>

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  Thin solid and dot-dashed curves
    correspond to our simple analytic estimates of $\dot{M}/\dot{M}_{\rm
      edd}$ (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively.  Thick curves correspond to the more accurate implicit analytic expression given by equation \eqref{eq:rs2main}.}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{cooling.eps}
  \caption{\label{fig:cooling} Ratio of the rates of heating to
radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$, as a
function of radius (solid lines) for each of our solutions shown in
Figure \ref{fig:profiles}.  Dashed lines show the ratio
of the cooling timescale to the free-fall timescale $t_{\rm cool}/t_{\rm ff}$.
For high heating rates both ratios are approximately equal at the
stagnation radius (squares).  When $\dot{q}_{\rm
heat}/\dot{q}_{\rm rad} \lesssim 10$ (or, equivalently, $t_{\rm cool}/t_{\rm ff}
\lesssim 10$ near the stagnation radius), then the flow is susceptible to thermal instabilities.}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{vwSourcesImp.pdf}
\caption{\label{fig:vwSourcesImp} Sources contributing to
  the gas heating rate at time $\tau_{\star}$ after
  a single burst of star formation.  Top and bottom panels show black hole masses of $\Mbh=10^6 \Msun$ and $\Mbh= 10^8 \Msun$
  (both cusps with $\Gamma=0.8$).  Solid and dashed lines show the ranges of
  $\tau_{\star}$ for which the accretion flow is thermally stable and
  unstable, respectively, according to the ratio of $\dot{q}_{\rm
    heat}/|\dot{q}_{\rm rad}|$ near the stagnation radius
  (eq.~[\ref{eq:cooling3}]).  Shown with horizontal gray lines
  are the stellar velocity dispersion for each SMBH mass, estimated as $\sigma_0=\sqrt{3} \sigma_{\bullet}$ (eq.~[\ref{eq:Msigma}]).}
\end{figure}

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of heating sources as a function of time $\tau_{\star}$ after a burst of
star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top) and $\Mbh = 10^{8} \Msun$ (bottom). The heating and mass
input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$
and $\eta(\tau_{\star})$, are calculated as described in
Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The SN Ia and Compton heating rates, $v_{w}^{\rm Ia}$ and $v_{\rm w}^{\rm C}$, are
calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC},
respectively, which also make use of $\eta(\tau_{\star})$ as set by stellar winds.\footnote{Both $v_{\rm w}^{\rm C}$ and
$v_{\rm w}^{\rm Ia}$ depend on the total heating rate
(eq.~[\ref{eq:vtot}]), requiring us to simultaneously solve a series
of implicit equations to determine each.}  SN Ia heating depends also on a convolution of the DTD (eq. [\ref{eq:DTD}]) and the SFH, which reduces to the DTD itself for a single star burst.  We account for the expected suppression of SN Ia heating resulting from its non-steady nature by setting $v_{w}^{\rm Ia} = 0$ when the stagnation radius (calculated {\it excluding} Ia heating) exceeds
the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]).  
Heating from

Figure \ref{fig:vwSourcesImp} shows that stellar winds are the most important heating source at early times, $\tau_{\star} \lesssim 10^{7}$ years.  Compton heating becomes more important at later times due to (1) the higher accretion rates that accompany the overall decrease in all sources of heating, coupled with (2) the persistently high mass loss rates and gas densities associated with the still relatively young stellar population.  Interestingly, for $M_{\bullet} = 10^{6}M_{\odot}$, SN Ia dominate the heating briefly around 10$^{9}$ yr due to a window of time in which $r_{\rm Ia} < \rs$.  For high-mass black holes (Fig.~\ref{fig:vwSourcesImp}, bottom panel), SN Ia dominate at all times after $10^{7}$ yr.

As the total heating rate declines with time, the flow inevitably becomes thermally unstable according to the criterion $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim 10$
(eq.~[\ref{eq:cooling3}]), as shown by dashed lines in
Fig.~\ref{fig:vwSourcesImp}.  Compton heating is neglected in
calculating thermal stability because$-$unlike local stellar feedback
mechanisms$-$it is not clear that SMBH feedback is capable of stabilizing the
flow given its inability to respond instantaneously to local changes
in gas properties.  Thermally unstable flow is present throughout the single starburst case except at very early times, $\tau_{\star} \lesssim 10^7$ yr, in the low $M_{\bullet}$ case and at late times $\tau_{\star}
\gtrsim 10^{9}$ yr in the high $M_{\bullet}$ case.

Finally, note that MSP heating is negligibly small for fiducial parameters and hence
is not shown, while kinetic feedback is neglected given its uncertain efficiency ($\S\ref{sec:kinetic}$).
